Light and growth form interact to shape stomatal ratio among British angiosperms

### 2017-10-04 ms/ms.Rnw
added figure comparing growth forms
modified table 1

Finally, phylogenetic information should improve inferences about paleoclimates because there is appreciable phylogenetic signal in stomatal ratio. The phylogenetic half-life of stomatal ratio evolution, after accounting for \el~and Raunki\ae r life form, is $\mathrm{log}(2) / \alpha = \Sexpr{round(log(2) / fitSRa$optpar, 1)}$ my (see Table~\ref{table:srAIC} for maximum likelihood estimates of $\alpha$, the return rate in the Ornstein-Uhlenbeck model of trait evolution). This lag time may indicate that evolving to the `optimum' is constrained by the shape of the fitness landscape \citep{Muir_2015} or that other unmeasured factors which affect stomatal ratio have some phylogenetic signal. Regardless of the mechanism, this fact means that researchers may be able to use data from closely related species to improve paleoenvironment reconstruction.

@@

\begin{table}[ht]
 \setstretch{\stretchy}
  \topcaption{Interaction beween species' Ellenberg light indicator value (\el) and Raunki\ae r lifeform predict stomatal ratio ($\mathrm{SR_{even}}$). I compared phylogenetic linear models using the Akaike Information Criterion (AIC), where AIC = $2k - 2 \mathrm{log}(\mathcal{L})$. $k$ is the number of model parameters and $\mathcal{L}$ is the model likelihood. Given a set of candidate models, the difference in AIC between a model and the lowest AIC ($\Delta$AIC) indicates the relative fit of competing models. The correlation coefficient $r^2$ is another indicator of model fit. $\alpha$ and $\sigma^2$ are the return rate and diffusion parameters of the Ornstein-Uhlenbeck model of trait evolution. }
  \begin{center}
  \begin{tabular}{@{} l c c c c c c c @{}}
  \toprule
  Model: $\mathrm{SR_{even}} \sim$ &  $\alpha$ & $\sigma^2$ &  $r^2$ & $k$ &  log($\mathcal{L}$) &  AIC &  $\Delta$AIC \\
  \midrule  
    \el~$\times$ lifeform
  & \Sexpr{round(fitSRa$optpar, 2)}
  & \Sexpr{round(fitSRa$sigma2, 3)}
  & \Sexpr{round(cor(fitSRa$fitted.values, fitSRa$fitted.values + fitSRa$residuals) ^ 2, 2)}
  & \Sexpr{length(coef(fitSRa)) + 2}
  & \Sexpr{round(fitSRa$logLik, 1)}
  & \Sexpr{round(fitSRa$aic, 1)}
  & \Sexpr{round(fitSRa$aic - min(aicSR), 1)} \\
    \el~+ lifeform
  & \Sexpr{round(fitSRb$optpar, 2)}
  & \Sexpr{round(fitSRb$sigma2, 3)}
  & \Sexpr{round(cor(fitSRb$fitted.values, fitSRb$fitted.values + fitSRb$residuals) ^ 2, 2)}
  & \Sexpr{length(coef(fitSRb)) + 2}
  & \Sexpr{round(fitSRb$logLik, 1)}
  & \Sexpr{round(fitSRb$aic, 1)}
  & \Sexpr{round(fitSRb$aic - min(aicSR), 1)} \\
    \el
  & \Sexpr{round(fitSRc$optpar, 2)}
  & \Sexpr{round(fitSRc$sigma2, 3)}
  & \Sexpr{round(cor(fitSRc$fitted.values, fitSRc$fitted.values + fitSRc$residuals) ^ 2, 2)}
  & \Sexpr{length(coef(fitSRc)) + 2}
  & \Sexpr{round(fitSRc$logLik, 1)}
  & \Sexpr{round(fitSRc$aic, 1)}
  & \Sexpr{round(fitSRc$aic - min(aicSR), 1)} \\
    lifeform
  & \Sexpr{round(fitSRd$optpar, 2)}
  & \Sexpr{round(fitSRd$sigma2, 3)}
  & \Sexpr{round(cor(fitSRd$fitted.values, fitSRd$fitted.values + fitSRd$residuals) ^ 2, 2)}
  & \Sexpr{length(coef(fitSRd)) + 2}
  & \Sexpr{round(fitSRd$logLik, 1)}
  & \Sexpr{round(fitSRd$aic, 1)}
  & \Sexpr{round(fitSRd$aic - min(aicSR), 1)} \\
    null
  & \Sexpr{round(fitSRe$optpar, 2)}
  & \Sexpr{round(fitSRe$sigma2, 3)}
  & \Sexpr{0}
  & \Sexpr{length(coef(fitSRe)) + 2}
  & \Sexpr{round(fitSRe$logLik, 1)}
  & \Sexpr{round(fitSRe$aic, 1)}
  & \Sexpr{round(fitSRe$aic - min(aicSR), 1)} \\
	\bottomrule
	\end{tabular}
	\label{table:srAIC}
\end{center}
\end{table}

@@

>>>>

Finally, phylogenetic information should improve inferences about paleoclimates because there is appreciable phylogenetic signal in stomatal ratio. The phylogenetic half-life of stomatal ratio evolution, after accounting for \el~and Raunki\ae r life form, is $\mathrm{log}(2) / \alpha = \Sexpr{round(log(2) / fitSR_lf_a$optpar, 1)}$ my (see Table~\ref{table:srAIC} for maximum likelihood estimates of $\alpha$, the return rate in the Ornstein-Uhlenbeck model of trait evolution). This lag time may indicate that evolving to the `optimum' is constrained by the shape of the fitness landscape \citep{Muir_2015} or that other unmeasured factors which affect stomatal ratio have some phylogenetic signal. Regardless of the mechanism, this fact means that researchers may be able to use data from closely related species to improve paleoenvironment reconstruction.

@@

\begin{table}[ht]
 \setstretch{\stretchy}
  \topcaption{Interaction beween species' Ellenberg light indicator value (\el) and Raunki\ae r lifeform predict stomatal ratio ($\mathrm{SR_{even}}$). I compared phylogenetic linear models using the Akaike Information Criterion (AIC), where AIC = $2k - 2 \mathrm{log}(\mathcal{L})$. $k$ is the number of model parameters and $\mathcal{L}$ is the model likelihood. Given a set of candidate models, the difference in AIC between a model and the lowest AIC ($\Delta$AIC) indicates the relative fit of competing models. The correlation coefficient $r^2$ is another indicator of model fit. $\alpha$ and $\sigma^2$ are the return rate and diffusion parameters of the Ornstein-Uhlenbeck model of trait evolution. }
  \begin{center}
  \begin{tabular}{@{} l c c c c c c c @{}}
  \toprule
  Model: $\mathrm{SR_{even}} \sim$ &  $\alpha$ & $\sigma^2$ &  $r^2$ & $k$ &  log($\mathcal{L}$) &  AIC &  $\Delta$AIC \\
  \midrule  
    \el~$\times$ Raunki\ae r lifeform
  & \Sexpr{round(fitSR_lf_a$optpar, 2)}
  & \Sexpr{round(fitSR_lf_a$sigma2, 3)}
  & \Sexpr{round(cor(fitSR_lf_a$fitted.values, fitSR_lf_a$fitted.values + fitSR_lf_a$residuals) ^ 2, 2)}
  & \Sexpr{length(coef(fitSR_lf_a)) + 2}
  & \Sexpr{round(fitSR_lf_a$logLik, 1)}
  & \Sexpr{round(fitSR_lf_a$aic, 1)}
  & \Sexpr{round(fitSR_lf_a$aic - min(aicSR), 1)} \\
    \el~$\times$ growth form
  & \Sexpr{round(fitSR_gf_f$optpar, 2)}
  & \Sexpr{round(fitSR_gf_f$sigma2, 3)}
  & \Sexpr{round(cor(fitSR_gf_f$fitted.values, fitSR_gf_f$fitted.values + fitSR_gf_f$residuals) ^ 2, 2)}
  & \Sexpr{length(coef(fitSR_gf_f)) + 2}
  & \Sexpr{round(fitSR_gf_f$logLik, 1)}
  & \Sexpr{round(fitSR_gf_f$aic, 1)}
  & \Sexpr{round(fitSR_gf_f$aic - min(aicSR), 1)} \\
    \el~+ Raunki\ae r lifeform
  & \Sexpr{round(fitSR_lf_b$optpar, 2)}
  & \Sexpr{round(fitSR_lf_b$sigma2, 3)}
  & \Sexpr{round(cor(fitSR_lf_b$fitted.values, fitSR_lf_b$fitted.values + fitSR_lf_b$residuals) ^ 2, 2)}
  & \Sexpr{length(coef(fitSR_lf_b)) + 2}
  & \Sexpr{round(fitSR_lf_b$logLik, 1)}
  & \Sexpr{round(fitSR_lf_b$aic, 1)}
  & \Sexpr{round(fitSR_lf_b$aic - min(aicSR), 1)} \\
     \el~+ growth form
  & \Sexpr{round(fitSR_gf_g$optpar, 2)}
  & \Sexpr{round(fitSR_gf_g$sigma2, 3)}
  & \Sexpr{round(cor(fitSR_gf_g$fitted.values, fitSR_gf_g$fitted.values + fitSR_gf_g$residuals) ^ 2, 2)}
  & \Sexpr{length(coef(fitSR_gf_g)) + 2}
  & \Sexpr{round(fitSR_gf_g$logLik, 1)}
  & \Sexpr{round(fitSR_gf_g$aic, 1)}
  & \Sexpr{round(fitSR_gf_g$aic - min(aicSR), 1)} \\
    Raunki\ae r lifeform
  & \Sexpr{round(fitSR_lf_d$optpar, 2)}
  & \Sexpr{round(fitSR_lf_d$sigma2, 3)}
  & \Sexpr{round(cor(fitSR_lf_d$fitted.values, fitSR_lf_d$fitted.values + fitSR_lf_d$residuals) ^ 2, 2)}
  & \Sexpr{length(coef(fitSR_lf_d)) + 2}
  & \Sexpr{round(fitSR_lf_d$logLik, 1)}
  & \Sexpr{round(fitSR_lf_d$aic, 1)}
  & \Sexpr{round(fitSR_lf_d$aic - min(aicSR), 1)} \\
    growth form
  & \Sexpr{round(fitSR_gf_h$optpar, 2)}
  & \Sexpr{round(fitSR_gf_h$sigma2, 3)}
  & \Sexpr{round(cor(fitSR_gf_h$fitted.values, fitSR_gf_h$fitted.values + fitSR_gf_h$residuals) ^ 2, 2)}
  & \Sexpr{length(coef(fitSR_gf_h)) + 2}
  & \Sexpr{round(fitSR_gf_h$logLik, 1)}
  & \Sexpr{round(fitSR_gf_h$aic, 1)}
  & \Sexpr{round(fitSR_gf_h$aic - min(aicSR), 1)} \\
    \el
  & \Sexpr{round(fitSRc$optpar, 2)}
  & \Sexpr{round(fitSRc$sigma2, 3)}
  & \Sexpr{round(cor(fitSRc$fitted.values, fitSRc$fitted.values + fitSRc$residuals) ^ 2, 2)}
  & \Sexpr{length(coef(fitSRc)) + 2}
  & \Sexpr{round(fitSRc$logLik, 1)}
  & \Sexpr{round(fitSRc$aic, 1)}
  & \Sexpr{round(fitSRc$aic - min(aicSR), 1)} \\
    null
  & \Sexpr{round(fitSRe$optpar, 2)}
  & \Sexpr{round(fitSRe$sigma2, 3)}
  & \Sexpr{0}
  & \Sexpr{length(coef(fitSRe)) + 2}
  & \Sexpr{round(fitSRe$logLik, 1)}
  & \Sexpr{round(fitSRe$aic, 1)}
  & \Sexpr{round(fitSRe$aic - min(aicSR), 1)} \\
	\bottomrule
	\end{tabular}
	\label{table:srAIC}
\end{center}
\end{table}

@@

\begin{figure}[ht]
\centerline{\includegraphics{figures/figureS_compare-growth-forms.pdf}}
\caption{CAPTION}
\label{fig:compare-growth-forms}
\end{figure}
